NanoMaLy

### 2015-11-18 paper/pldi16-submission/related.tex
working on the related work

\subsection{Type Error Localization}
\label{sec:type-error-local}
\begin{itemize}
\item \cite{lerner_searching_2007}
\item \cite{zhang_toward_2014}
\item \cite{pavlinovic_finding_2014}
\item \cite{chen_counter-factual_2014}
\end{itemize}

\subsection{Testing}
\label{sec:testing}
\begin{itemize}
\item \cite{claessen_quickcheck:_2000}
\item \cite{runciman_smallcheck_2008}
\item \cite{godefroid_dart:_2005}
\item \cite{cadar_klee:_2008}
\item \cite{tillmann_pexwhite_2008}
\end{itemize}

>>>>

\subsection{Type Error Localization}
\label{sec:type-error-local}
It is well known that unification-based type inference procedures can
produce poor error messages, and in particular, can misidentify the
\emph{source} of the type error.
%

\begin{itemize}
\item Lerner \etal~\cite{lerner_searching_2007} attempt to localize the
  type error and suggest a fix by replacing expressions (or removing
  them entirely) with alternatives based on the surrounding program
  context.
\item Chen and Erwig~\cite{chen_counter-factual_2014} use a variational
  type system represent the possibility that the inferred type of an
  expression may be incorrect, because the expression may be the source
  of the error. It then attempts to deduce the error source by searching
  for an expression whose type can be changed such that type inference
  would succeed. In contrast to Seminal, which searches for changes at
  the value-level, by searching at the type level the search is complete
  due the finite universe of types in a program.
\item Neubauer and Thiemann~\cite{neubauer_discriminative_2003} present
  a decidable type system based on discriminative sum types, in which
  all terms are typeable and type derivations contain all type errors in
  a program. They then use the typing derivation to slice out the parts
  of the expression related to each error.
\item Zhang and Myers~\cite{zhang_toward_2014} present an algorithm for
  identifying the most likely culprit in a system of unsatisfiable
  constraints (\eg type equalities), based on Bayesian reasoning.
\item Pavlinovic \etal~\cite{pavlinovic_finding_2014} translate the
  error localization problem to a MaxSMT optimization problem, using
  compiler-provided weights to rank the possible sources.
% \item \cite{chen_error-tolerant_2012}
% \item \cite{okeefe_type_1992}
% \item \cite{gomard_partial_1990}
% \item \cite{thatte_type_1988}
\end{itemize}

In contrast to these approaches, we do not attempt to localize or fix
the type error. Instead we try to explain it to the novice user using a
dynamic witness that demonstrates how their program is not just
ill-typed but truly wrong. In addition, allowing users to run their
program (even knowing that it is wrong) enables experimentation and the
use of debuggers to step through the program and investigate its
evolution.

\subsection{Testing}
\label{sec:testing}
\nanomaly is at its heart a test generator, and builds on a rich line of
work.
%
Our use of holes to represent unknown values is inspired by the work of
Runciman, Naylor, and Lindblad~\cite{runciman_smallcheck_2008,naylor_finding_2007,lindblad_property_2007},
%
who use lazy evaluation to drastically reduce the search space for
exhaustive test generation, by grouping together equivalent inputs by
the set of values they force. An exhaustive search is complete (up to
the depth bound), if a witness exists it will be found, but due to the
exponential blowup in the search space the depth bound can be quite
limited without advanced grouping and filtering techniques.
%
Our search is not exhaustive; instead we use random generation to fill
in holes on demand.
%
Random test generation~\cite{claessen_quickcheck:_2000,csallner_jcrasher:_2004,pacheco_feedback-directed_2007}
%
is by its nature incomplete, but is able to check larger inputs than
exhaustive testing as a result.

Instead of enumerating values, which may trigger the same path through
the program, one might enumerate paths. 
%
Dynamic-symbolic execution~\cite{godefroid_dart:_2005,cadar_klee:_2008,tillmann_pex_2008}
%
enables this by symbolic execution (to track which path a given input
vector triggers) with concrete execution (to ensure test failures are
not spurious). The system collects a path condition during execution,
which tracks symbolically what conditions must be met to trigger the
current path. Upon successfully completing a test run, it negates the
path condition and queries a solver for another set of inputs that
satisfy the negated path condition, \ie inputs that will not trigger the
same path. Thus, dynamic-symbolic execution can prune the search space
much faster than techniques based on enumerating values, but is limited
by the expressiveness of the underlying solver. Our operational
semantics is amenable to dynamic-symbolic execution, one would just need
to collect the path condition and replace our implementation of \gensym
by a call to the solver. We chose to use lazy, random generation instead
because it does not incur the overhead of an external solver, and
produces high coverage for our domain of novice programs.

% \begin{itemize}
% \item \cite{claessen_quickcheck:_2000}
% \item 
% \item \cite{godefroid_dart:_2005}
% \item \cite{cadar_klee:_2008}
% \item \cite{tillmann_pex_2008}
% \end{itemize}

\subsection{Misc}
\begin{itemize}
\item \cite{vytiniotis_equality_2012} extend the Haskell compiler GHC to
  support compiling ill-typed programs, but their intent is rather
  different from ours. Their goal was to allow programmers to
  incrementally test refactorings, which often cause type errors in
  distant pieces of code. They replace any expression that fails to
  type check with a \emph{runtime} error, but do not perform
  any type checking at runtime.
\item \cite{perera_functional_2012}
\end{itemize}
